---
title: "RawHumms Data Quality Control Report"
date: '`r format(Sys.Date(), "%B, %Y")`'
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
params:
  msdata: NA
  mypeaks: NA
  myppm: NA
  myrt: NA
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
# get data and clean them
msdata <- params$msdata
mypeaks <- params$mypeaks
myppm <- params$myppm
myrt <- params$myrt
## remove rows with m/z being NA in mypeaks and rename RT column
mypeaks <- as.data.frame(mypeaks) %>% 
  rename(rt = Expected_RT) %>%
  filter(!is.na(mz))
```

## Data Overview

Below is an interactive total ion current (TIC) plot.

```{r, echo = FALSE, out.width = "100%"}
plotTIC <- ggplot(msdata$TIC) + 
  geom_line(aes(x = rt, y = int, color = filename)) +
  facet_wrap(~filename, scales = "free_y", ncol = 1) +
  labs(x = "Retention Time (min)", y = "Intensity", color = "File name: ") +
  theme(legend.position = "top") +
  theme_bw()

ggplotly(plotTIC)
```

Here is the EIC

```{r, echo = FALSE, out.width = "100%"}
plotEIC <- ggplot(msdata$BPC) + 
  geom_line(aes(x = rt, y = int, color = filename)) +
  facet_wrap(~ filename, scales = "free_y", ncol = 1) +
  labs(x = "Retention Time (min)", y = "Intensity", color = "File name: ") +
  theme(legend.position = "top") +
  theme_bw()

ggplotly(plotEIC)
```

Below is the ion intensity plot. Dash lines represents the mean peak area ± 1.96 SD.

```{r, echo = FALSE, warning = FALSE, message = FALSE, out.width = "100%"}
sumTIC <- msdata$TIC %>%
  group_by(filename) %>%
  summarise(peakArea = sum(int)) %>%
  mutate(Mean = mean(peakArea),
         SD = sd(peakArea),
         CV = round(SD/Mean*100, 2))

plotSumTIC <- ggplot(sumTIC, aes(x = filename, y = peakArea, group = 1)) +
  geom_bar(stat = "identity", aes(fill = filename), alpha = 0.5) +
  theme_bw()

ggplotly(plotSumTIC)
```


```{r cars}
head(mypeaks)
print(myppm)
print(myrt)
```

## MS1

`RawHummus` automatically search for **6** peaks across the RT, and evaluate them.

```{r echo = FALSE, out.width = "100%"}
# get min and max RT of madata
minRT <- min(msdata$MS1$rt)
maxRT <- max(msdata$MS1$rt)
# get lists of peaks for evaluation
topPeaks <- msdata$MS1 %>%
  mutate(bins = ntile(rt, 6)) %>%
  group_by(bins) %>%
  slice_max(int, n = 1) %>%
  ungroup() %>%
  select(rt, mz)
# search in raw data
autoTarget <- vector(mode = "list", length = dim(topPeaks)[1])
for (i in 1:dim(topPeaks)[1]){
  autoTarget[[i]] <- msdata$MS1[mz%between%pmppm(topPeaks$mz[i], myppm) & 
                                 rt%between%c(max(topPeaks$rt[i] - myrt, minRT), min(topPeaks$rt[i] + myrt,maxRT))]
  names(autoTarget)[i] <- paste0("mz:", round(topPeaks$mz[i], 3), " RT:", round(topPeaks$rt[i], 2))
}
# unlist
autoTarget <- dplyr::bind_rows(autoTarget, .id = "id")
# plot
if (nrow(autoTarget) > 0) {
    pp <- ggplot(autoTarget) + 
      geom_line(aes(x = rt, y = int, color = filename)) +
      facet_wrap(~ id, scales = "free_y", ncol = 2) +
      labs(x = "Retention time (min)", y = "Intensity", color = "File name: ") +
      theme(legend.position = "top") +
      theme_bw()
    ggplotly(pp)
    } else {
      print("Peasks of interests are not found")
    }

```

You can also embed plots, for example:

```{r echo = FALSE, out.width = "100%"}
if(!all(is.na(mypeaks))) {
  # search according to mz and RT with tolerances
  getTarget <- vector(mode = "list", length = dim(mypeaks)[1])
  for (i in 1:dim(mypeaks)[1]){
  getTarget[[i]] <- msdata$MS1[mz%between%pmppm(mypeaks$mz[i], myppm) & 
                                 rt%between%c(max(mypeaks$rt[i] - myrt, minRT), min(mypeaks$rt[i] + myrt,maxRT))]
  names(getTarget)[i] <- paste0("mz:", round(mypeaks$mz[i], 4), " RT:", round(mypeaks$rt[i], 2))
  }
  
  # unlist
  myTarget <- dplyr::bind_rows(getTarget, .id = "id")
  # plot
  if (nrow(myTarget) > 0) {
    pp <- ggplot(myTarget) + 
      geom_line(aes(x = rt, y = int, color = filename)) +
      facet_wrap(~ id, scales = "free_y", ncol = 2) +
      labs(x = "Retention time (min)", y = "Intensity", color = "File name: ") +
      theme(legend.position = "top") +
      theme_bw()
    ggplotly(pp)
    } else {
      print("Peasks of interests are not found")
    }
} else {
    print("Peaks of interests are not defined")
  }
```

```{r}
head(mypeaks)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
