---
title: "RawHumms Data Quality Control Report"
date: '`r format(Sys.Date(),  "%d, %B, %Y")`'
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    number_sections: true
    css: style.css
params:
  msdata: NA
  mynoise: NA
  mypeaks: NA
  myppm: NA
  myrt: NA
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# get data and clean them
msdata <- params$msdata
mypeaks <- params$mypeaks
mynoise <- params$mynoise
myppm <- params$myppm
myrt <- params$myrt
## remove rows with m/z being NA in mypeaks and rename RT column
mypeaks <- as.data.frame(mypeaks) %>% 
  rename(rt = Expected_RT) %>%
  filter(!is.na(mz))

```

<img src = "logo.png" height = "120" style = "position: absolute; top: 20px; right: 10px;"/>

<br>

# Introduction

Robust and reproducible data is essential to ensure confidence in analytical results and is particularly important for large-scale metabolomics studies. Therefore raw data need to be inspected before data processing and statistical analysis in order to detect measurement bias and verify system consistency. In liquid chromatography mass spectrometry (LCMS) based metabolomics studies, proper quality control (QC) checks are particularly important to ensure reliable and comparable results within experimental measurements [1-2].

<span style = "color : #FF8300">**RawHummus**</span> is an user-friendly web application for rapid data quality check based on raw QC samples. It generates an HTML report with interactive plots, statistics and illustrations that help users evaluate their data quality and LCMS system performance.

# Chromatogram

## TIC plot

`Total Ion Current (TIC)` chromatogram represents the summed ion intensity across each scan in the analysis. **The interactive overlaid TIC plot** can be used for rapid inspection of retention time (RT) and ion intensity fluctuations.

```{r TIC, echo = FALSE, out.width = "100%"}

plotTIC <- ggplot(msdata$TIC) + 
  geom_line(aes(x = rt, y = int, color = filename)) +
  labs(x = "RT (min)", y = "Intensity", color = "File name: ") +
  theme_bw()

ggplotly(plotTIC)

```

## Summed TIC Plot

**Summed TIC plot** is another quick-and-dirty way to overview global ion intensity variations among QC samples. It summed TIC across the entire points (scans) from the analysis. <span style = "color: #ff8080">**Dashed red line**</span> is **mean** of summed TIC and <span style = "color: #0579fc">**blue lines**</span> represent **mean + 2SD** and **mean - 2SD**, respectively.

```{r sumTIC, echo = FALSE, warning = FALSE, message = FALSE, out.width = "100%"}

sumTIC <- msdata$TIC %>% 
  group_by(filename) %>%
  summarise(peakArea = sum(int)) %>%
  mutate(Mean = mean(peakArea),
         "Mean+2SD" = mean(peakArea) + 2*sd(peakArea),
         "Mean-2SD" = mean(peakArea) - 2*sd(peakArea),)

plotSumTIC <- ggplot(sumTIC, aes(x = filename, y = peakArea, group = 1)) +
  geom_bar(stat = "identity", aes(fill = filename), alpha = 0.5) +
  geom_hline(aes(yintercept = Mean), linetype = 'dotted', colour = "red", alpha = 0.5) +
  geom_hline(aes(yintercept = `Mean+2SD`), colour = "blue", linetype = 'dotted', alpha = 0.5) +
  geom_hline(aes(yintercept = `Mean-2SD`), colour = "blue", linetype = 'dotted', alpha = 0.5) +
  labs(x = "File name", y = "Summed TIC", fill = "File name") +
  theme_bw() +
  theme(axis.text.x = element_blank())

ggplotly(plotSumTIC)

```

## TIC Correlation Analysis

`Pearson correlation` is  used to quantify the metabolic profile similarity among QC samples. Pearson correlation coefficient (**R**) over **0.85** indicate high metabolic profile similarity in RT and chromatogram peak shape. If the value is below <span style = "color: #ff8080">**0.85**</span>, it will be highlight in <span style = "color: #ff8080">**red**</span> in Table 1. 

> **Note that** RT were binned by 0.1 min for Pearson correlation calculation.

```{r corTIC, echo = FALSE, warning = FALSE, message = FALSE, out.width = "100%"}

## if scan No. is the same, corTIC is a DF, otherwise it is a list (this case is not common after RT binning)

corTIC <- msdata$TIC %>%
  filter(rt >= 0.5) %>% # trim first 0.5 min to avoid unequal scans at the beginning
  mutate(rt_bins = cut(rt, breaks = floor(max(rt)/0.1))) %>% # bin RT 0.1 min, it is important to bin raw data
  group_by(rt_bins, filename) %>%
  mutate(int2 = sum(int), rt2 = mean(rt)) %>%
  ungroup() %>%
  select(filename, int2, rt2) %>%
  distinct_all() %>%
  select(int2, filename) %>%
  unstack()

## calculate correlation.
if(is.data.frame(corTIC)){
  as.data.frame(round(cor(corTIC), 3)) %>%
    dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85, "#ff8080", "#FFFFFF"))) %>%
    kableExtra::kbl(escape = FALSE, caption = "Table 1: Metabolic profile similarity") %>%
    kableExtra::kable_classic("hover", full_width = TRUE) %>%
    kableExtra::scroll_box(width = "100%")
} else {
  ## get the minimum scan number
  minScan <- min(map_df(corTIC, length))
  
  ## trim scans
  corTIC2 <- map_df(corTIC, `[`, 1: minScan)
  
  ## correlation table
  as.data.frame(round(cor(corTIC2), 2)) %>%
    dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85 , "#ff8080", "#FFFFFF"))) %>%
    kableExtra::kbl(escape = FALSE, 
                    caption = paste0("Table 1: Metabolic profile similarity (Sample trimmed to ", minScan, " TIC scans)")) %>%
    kableExtra::kable_classic("hover", full_width = TRUE) %>%
    kableExtra::scroll_box(width = "100%")
}

```

> **Note that** `TIC Correlation Analysis` is mainly used to evaluate peak shape similarity and RT consistency. QC files with ion intensity drift but similar profiles could still have good Pearson correlation coefficient.


# MS1

## Auto Peaks Evaluation

In order to accurately monitor variations in mass, RT and ion intensity, <span style = "color : #FF8300">**RawHummus**</span> automatically selects **6** peaks across the entire RT range, and use them to evaluate LCMS system.

Below are the Extracted ion chromatogram (EIC) of the 6 selected ions. You can interactively view, inspect and compare them among different QC samples.

```{r MS1Plot, echo = FALSE, out.width = "100%"}

# get lists of peaks for evaluation

MS1Filter <- msdata$MS1[int > mynoise] ## filter according user defined noise

# get min and max RT of msdata
minRT <- min(MS1Filter$rt)
maxRT <- max(MS1Filter$rt)

##############################################################################
# topPeaks <- MS1Filter %>% 
#   mutate(bins = as.factor(cut(rt, 6, labels = FALSE))) %>%
#   group_by(bins) %>%
#   slice_max(int, n = 1, with_ties = FALSE) %>%
#   ungroup() %>%
#   select(rt, mz)
##############################################################################

## use data.table instead of dplyr
topPeaks <- MS1Filter[, bins := cut(rt, 6, labels = FALSE)] %>%
  .[, .SD[which.max(int)], by = bins] %>%
  .[, list(rt, mz)]

# search in raw data
autoTarget <- vector(mode = "list", length = dim(topPeaks)[1])

##############################################################################
# for (i in 1:dim(topPeaks)[1]){
#   autoTarget[[i]] <- MS1Filter %>%
#     filter(between(mz, pmppm(topPeaks$mz[i], myppm)[1], pmppm(topPeaks$mz[i], myppm)[2])) %>%
#     filter(between(rt, max(topPeaks$rt[i] - myrt, minRT), min(topPeaks$rt[i] + myrt, maxRT)))
#   names(autoTarget)[i] <- paste0("RT: ", round(topPeaks$rt[i], 2), " mz: ", round(topPeaks$mz[i], 3))
# }
##############################################################################

## use data.table instead of dplyr
for (i in 1:dim(topPeaks)[1]){
  autoTarget[[i]] <- MS1Filter[mz %between% pmppm(topPeaks$mz[i]) & rt %between% c(max(topPeaks$rt[i] - myrt, minRT), min(topPeaks$rt[i] + myrt, maxRT))]
  names(autoTarget)[i] <- paste0("RT: ", round(topPeaks$rt[i], 2), " mz: ", round(topPeaks$mz[i], 3))
}


##############################################################################
# unlist
# autoTarget <- dplyr::bind_rows(autoTarget, .id = "id")
##############################################################################

## unlist
autoTarget <- data.table::rbindlist(autoTarget, idcol = T)
data.table::setnames(autoTarget, ".id", "id") ## keep the same name as in dplyr
autoTarget <- autoTarget %>%
  mutate(id = factor(id)) %>%
  mutate(id = factor(id, levels = levels(id)[order(read.table(text = levels(id), fill = TRUE)[[2]])])) # order by RT

# plot
if (nrow(autoTarget) > 0) {
    aPlot <- ggplot(autoTarget) + 
      geom_line(aes(x = rt, y = int, color = filename)) +
      facet_wrap(~ id, scales = "free_y", ncol = 2) +
      labs(x = "RT (min)", y = "Intensity", color = "File name: ") +
      theme_bw()
    ggplotly(aPlot)
    } else {
      print("Peasks of interests are not found")
    }

```

<span style = "color: #FF8300">**RawHummus**</span> performs a simple statistics to make rapid evaluation. The table below summarized the comparison result, in which maximum RT difference, mass difference and ion intensity difference, and Intensity CV are given.

`Max RT Diff (min)`: is the maximum retention time variation (in min unit). Small value indicates a good retention time consistency. If the maximum retention time variation is over <span style = "color: #ff8080">**1 min**</span>, the value will be highlight in <span style = "color: #ff8080">**red**</span> in Table 2. 

`Max Mass Diff (ppm)`: is the maximum mass variation (in ppm unit). Small values indicate good mass accuracy. If the maximum mass variation is over <span style = "color: #ff8080">**5 ppm**</span>, the value will be highlight in <span style = "color: #ff8080">**red**</span> in Table 2.

`Max Intensity Fold Change`: is the maximum ion intensity variation. The value closer to 1 suggests that the ion intensity is stable. If the maximum intensity ratio is over <span style = "color: #ff8080">**2**</span>, the value will be highlight in <span style = "color: #ff8080">**red**</span> in Table 2.

If a peak is missing in some of your samples, <span style = "color: #ff8080">**NA**</span> values will be give in the table. You need to carefully inspect the peak so as to evaluate the reproducibility.

`Intensity CV (%)`: is the ion intensity coefficient of variation (also termed as relative standard deviation, STD). Smaller value indicates better ion intensity consistency. If the intensity CV is over <span style = "color: #ff8080">**30%**</span>, the value will be highlight in <span style = "color: #ff8080">**red**</span> in Table 2.

```{r MS1Table, echo = FALSE, out.width = "100%"}

if (nrow(autoTarget) > 0) {
  
##############################################################################
 # autoTargetTable <- autoTarget %>%
 #  group_by(id, filename) %>%
 #  slice_max(int) %>%
 #  tidyr::complete(id, filename) %>%
 #  group_by(id) %>%
 #  mutate("Max RT Diff (min)" = round(max(rt) - min(rt), 2),
 #         "Max Mass Diff (ppm)" = round((max(mz) - min(mz))/min(mz) * 10^6, 2),
 #         "Max Intensity Ratio" = round(max(int) / min(int), 2)) %>%
 #  mutate(Peak = id) %>%
 #  ungroup() %>%
 #  select(Peak, "Max RT Diff (min)", "Max Mass Diff (ppm)", "Max Intensity Ratio") %>%
 #  distinct_all() %>% 
 #  mutate("Max RT Diff (min)" = kableExtra::cell_spec(`Max RT Diff (min)`,
 #                                                     background = ifelse(`Max RT Diff (min)` > 1, "#ff8080", "#FFFFFF")),
 #         "Max Mass Diff (ppm)" = kableExtra::cell_spec(`Max Mass Diff (ppm)`, 
 #                                                       background = ifelse(`Max Mass Diff (ppm)` > 5, "#ff8080", "#FFFFFF")),
 #         "Max Intensity Ratio" = kableExtra::cell_spec(`Max Intensity Ratio`, 
 #                                                       background = ifelse(`Max Intensity Ratio` > 3, "#ff8080", "#FFFFFF")))
##############################################################################
  
  ## fix bug#2
  autoTargetTable  <- autoTarget[, .SD[which.max(int)], by = .(id, filename)] %>% 
  ## data.table fill missing rows 
  .[CJ(id = unique(id), filename = unique(msdata$TIC$filename)), on = c('id','filename')] %>%
  .[, `:=`("Max RT Diff (min)" =  round(max(rt) - min(rt), 2),
           "Max Mass Diff (ppm)" = round((max(mz) - min(mz))/min(mz) * 10^6, 2),
           "Max Intensity Ratio" = round(max(int) / min(int), 2),
           "Intensity CV (%)" = round(sd(int)/mean(int) * 100, 2)), by = id] %>%
  mutate(Peak = id) %>%
  select(Peak, "Max RT Diff (min)", "Max Mass Diff (ppm)", "Max Intensity Ratio", "Intensity CV (%)") %>%
  distinct_all() %>%
  mutate("Max RT Diff (min)" = kableExtra::cell_spec(`Max RT Diff (min)`, 
                                                     background = ifelse((`Max RT Diff (min)` > 1 | is.na(`Max RT Diff (min)`)) , "#ff8080", "#FFFFFF")),
         "Max Mass Diff (ppm)" = kableExtra::cell_spec(`Max Mass Diff (ppm)`, 
                                                       background = ifelse((`Max Mass Diff (ppm)` > 5 | is.na(`Max Mass Diff (ppm)` )), "#ff8080", "#FFFFFF")),
         "Max Intensity Ratio" = kableExtra::cell_spec(`Max Intensity Ratio`, 
                                                       background = ifelse(c(`Max Intensity Ratio` > 2 | is.na(`Max Intensity Ratio`)), "#ff8080", "#FFFFFF")),
         "Intensity CV (%)" = kableExtra::cell_spec(`Intensity CV (%)`, 
                                                       background = ifelse(c(`Intensity CV (%)` > 30 | is.na(`Intensity CV (%)`)), "#ff8080", "#FFFFFF")))
 
 ## use kableExtra instead of DT
 kableExtra::kbl(autoTargetTable, escape = FALSE, caption = "Table 2: Summary of auto-selected peaks") %>%
  kableExtra::kable_classic("hover", full_width = TRUE)

} else {
  print("Peasks of interests are not found")
}

```

## User defined peaks

Additionally, users could add their peaks of interests for inspection and comparison. If these peaks are defined in <span style = "color: #FF8300">**RawHummus**</span> and are found in the data. Similar plots and a data summary table will be given below. Otherwise, this section will be left blank.

> **Note that** noise peaks can be used to minitor the mass accuracy variation, but they may not work well to evalute RT and ion intensity variation.

```{r userMS1Plot, echo = FALSE, out.width = "100%"}

if(!all(is.na(mypeaks))) {
  
  # search according to mz and RT with tolerances
  getTarget <- vector(mode = "list", length = dim(mypeaks)[1])
  
  ##############################################################################
  # for (i in 1:dim(mypeaks)[1]){
  #   getTarget[[i]] <- MS1Filter %>%
  #     filter(between(mz, pmppm(mypeaks$mz[i], myppm)[1], pmppm(mypeaks$mz[i], myppm)[2])) %>%
  #     filter(between(rt, max(mypeaks$rt[i] - myrt, minRT, na.rm = TRUE), min(mypeaks$rt[i] + myrt, maxRT, na.rm = TRUE)))
  #   names(getTarget)[i] <- paste0(" RT: ", round(mypeaks$rt[i], 2), " mz: ", round(mypeaks$mz[i], 3))
  # }
  
  # unlist
  # myTarget <- dplyr::bind_rows(getTarget, .id = "id")
  ##############################################################################
  
  ## use data.table instead of dplyr
  for (i in 1:dim(mypeaks)[1]){
  getTarget[[i]] <- MS1Filter[mz %between% pmppm(mypeaks$mz[i]) & 
                                rt %between% c(max(mypeaks$rt[i] - myrt, minRT, na.rm = TRUE), min(mypeaks$rt[i] + myrt, maxRT, na.rm = TRUE))]
  names(getTarget)[i] <- paste0(" RT: ", round(mypeaks$rt[i], 2), " mz: ", round(mypeaks$mz[i], 3))
  }
  
  # unlist
  myTarget <- data.table::rbindlist(getTarget, idcol = T)
  data.table::setnames(myTarget, ".id", "id") ## keep the same name as in dplyr
  
  # plot
  if (nrow(myTarget) > 0) {
    uPlot <- ggplot(myTarget) + 
      geom_line(aes(x = rt, y = int, color = filename)) +
      facet_wrap(~ id, scales = "free_y", ncol = 2) +
      labs(x = "RT (min)", y = "Intensity", color = "File name: ") +
      theme_bw()
    ggplotly(uPlot)
    }
  
} else {
  myTarget <- NULL
  print("User defined peaks are not found")
}

```


```{r userMS1Table, echo = FALSE, out.width = "100%"}

if (!is.null(myTarget)) {
  
   ##############################################################################
  
 # myTargetTable <- myTarget %>%
 #  group_by(id, filename) %>%
 #  slice_max(int) %>%
 #  tidyr::complete(id, filename) %>%
 #  group_by(id) %>%
 #  mutate("Max RT Diff (min)" = round(max(rt) - min(rt), 2),
 #         "Max Mass Diff (ppm)" = round((max(mz) - min(mz))/min(mz) * 10^6, 2),
 #         "Max Intensity Ratio" = round(max(int) / min(int), 2)) %>%
 #  mutate(Peak = id) %>%
 #  ungroup() %>%
 #  select(Peak, "Max RT Diff (min)", "Max Mass Diff (ppm)", "Max Intensity Ratio") %>%
 #  distinct_all() %>%
 #  mutate("Max RT Diff (min)" = kableExtra::cell_spec(`Max RT Diff (min)`,
 #                                                     background = ifelse(`Max RT Diff (min)` > 1, "#ff8080", "#FFFFFF")),
 #         "Max Mass Diff (ppm)" = kableExtra::cell_spec(`Max Mass Diff (ppm)`, 
 #                                                       background = ifelse(`Max Mass Diff (ppm)` > 5, "#ff8080", "#FFFFFF")),
 #         "Max Intensity Ratio" = kableExtra::cell_spec(`Max Intensity Ratio`, 
 #                                                       background = ifelse(`Max Intensity Ratio` > 3, "#ff8080", "#FFFFFF")))
  
   ##############################################################################
  
   myTargetTable  <- myTarget[, .SD[which.max(int)], by = .(id, filename)] %>% 
  ## data.table fill missing rows 
     .[CJ(id = unique(id), filename = unique(msdata$MS1$filename)), on = c('id','filename')] %>%
     .[, `:=`("Max RT Diff (min)" =  round(max(rt) - min(rt), 2),
              "Max Mass Diff (ppm)" = round((max(mz) - min(mz)) / min(mz) * 10^6, 2),
              "Max Intensity Ratio" = round(max(int) / min(int), 2),
              "Intensity CV (%)" = round(sd(int) / mean(int) * 100, 2)), by = id] %>%
     mutate(Peak = id) %>%
     select(Peak, "Max RT Diff (min)", "Max Mass Diff (ppm)", "Max Intensity Ratio", "Intensity CV (%)") %>%
     distinct_all() %>%
     mutate("Max RT Diff (min)" = kableExtra::cell_spec(`Max RT Diff (min)`,
                                                        background = ifelse((`Max RT Diff (min)` > 1 | is.na(`Max RT Diff (min)`)) , "#ff8080", "#FFFFFF")),
            "Max Mass Diff (ppm)" = kableExtra::cell_spec(`Max Mass Diff (ppm)`, 
                                                        background = ifelse((`Max Mass Diff (ppm)` > 5 | is.na(`Max Mass Diff (ppm)` )), "#ff8080", "#FFFFFF")),
            "Max Intensity Ratio" = kableExtra::cell_spec(`Max Intensity Ratio`, 
                                                        background = ifelse(c(`Max Intensity Ratio` > 3 | is.na(`Max Intensity Ratio`)), "#ff8080", "#FFFFFF")),
            "Intensity CV (%)" = kableExtra::cell_spec(`Intensity CV (%)`, 
                                                        background = ifelse(c(`Intensity CV (%)` > 30 | is.na(`Intensity CV (%)`)), "#ff8080", "#FFFFFF")))

## use kableExtra instead of DT 
 kableExtra::kbl(myTargetTable, escape = FALSE, caption = "Table 3: Summary of user-defined peaks") %>%
  kableExtra::kable_classic("hover", full_width = TRUE)
 
 }

```

# MS2

MS/MS fragmentation is important for metabolite identification. RawHummus is also able to identify problems with regard to fragmentation.

> **Note that** if your data files do not contain any MS/MS information, this section will be left blank.

## Number of MS2 Events

`Number of triggered MS/MS`: is the total number of MS/MS events in the sample. Similar number of triggered MS/MS events indicates good reproducibility.


```{r MS2, echo = FALSE, warning = FALSE}

if(nrow(msdata$MS2) == 0){
  getMS2 <- NULL
  MS2_mz <- NULL
  MS2_RT <- NULL
  TableMS2Pmass <- NULL
  TableMS2PRT <- NULL
} else {
  getMS2 <- msdata$MS2 %>% 
    group_by(filename) %>%
    distinct(rt, premz) %>%
    summarise("MS2 Events" = n())
  
  MS2unique <- msdata$MS2 %>%
    distinct(rt, premz, voltage, filename)
  
  ## define cosine similarity function
  cosSim <- function(X){
  X2 <- as.matrix(X)
  denom <- solve(diag(sqrt(diag(t(X2)%*%X2))))
  Sim <- as.data.frame(denom%*%(t(X2)%*%X2)%*%denom)
  colnames(Sim) <- colnames(X)
  rownames(Sim) <- colnames(X)
  return(Sim)
  } 
  
  ##1.1 plot precursor ion across mz range. Split subplot for better visulization for large data files.
  ## here nsub is used to better split the plot, 8 is the max plot in one subplot
  nfile <- length(levels(as.factor(MS2unique$filename)))
  nsub <- ceiling(nfile/ceiling(nfile/8))

  MS2_mz <- MS2unique %>% 
  split(ceiling(group_indices(., filename)/nsub)) %>% 
  map(~ggplot(., aes(premz)) +
        geom_density(aes(fill = filename), alpha = 0.3) +
        facet_wrap(~ filename, scales = "free_y", ncol = 2) + 
        labs(x = "Mass Range", y = "Density", fill = "File name") +
        theme_bw() +
        theme(legend.position = "none"))
  
  ##1.2 for MS2_mz table
  corPmass <- MS2unique %>%
  mutate(mz_bins = cut(premz, breaks = floor(max(premz)/1))) %>% ## bin by 10 Da
  group_by(filename) %>%
  count(mz_bins) %>%
  select(n, filename) %>%
  unstack()
  
  if(is.data.frame(corPmass)){
  TableMS2Pmass <- as.data.frame(round(ccosSim(corPmass), 3)) %>%
    dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85 , "#ff8080", "#FFFFFF"))) %>%
    kableExtra::kbl(escape = FALSE, caption = "Table 5: Precursor Distribution across mass similarity") %>%
    kableExtra::kable_classic("hover", full_width = TRUE) %>%
    kableExtra::scroll_box(width = "100%")
    } else {
      ## get the minimum scan number
      minScan <- min(map_df(corPmass, length))
      ## trim scans
      corPmass2 <- map_df(corPmass, `[`, 1: minScan)
      ## correlation table
      TableMS2Pmass <- as.data.frame(round(cosSim(corPmass2), 3)) %>%
        dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85 , "#ff8080", "#FFFFFF"))) %>%
        kableExtra::kbl(escape = FALSE, 
                    caption = "Table 5: Precursor Distribution across mass similarity") %>%
        kableExtra::kable_classic("hover", full_width = TRUE) %>%
        kableExtra::scroll_box(width = "100%")
      }

  
  ##2.1 plot precursor ions across RT range. Split subplot for better visualization of large data files.
  MS2_RT <- MS2unique %>% 
  split(ceiling(group_indices(., filename)/nsub)) %>% 
  map(~ggplot(., aes(rt, fill = filename)) +
        geom_density(alpha = 0.3) +
        facet_wrap(~ filename, scales = "free_y", ncol = 2) + 
        labs(x = "RT (min)", y = "Density", fill = "File name") +
        theme_bw() +
        theme(legend.position = "none"))
  
  ##2.2 for MS2_RT table
  corPRT <- MS2unique %>%
    mutate(rt_bins = cut(rt, breaks = floor(max(rt)/0.05))) %>% ## bin by 0.05 min
    group_by(filename) %>%
    count(rt_bins) %>%
    select(n, filename) %>%
    unstack()
  
  if(is.data.frame(corPRT)){
  TableMS2PRT <- as.data.frame(round(cosSim(corPRT), 3)) %>%
    dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85 , "#ff8080", "#FFFFFF"))) %>%
    kableExtra::kbl(escape = FALSE, caption = "Table 6: Precursor Distribution across RT similarity") %>%
    kableExtra::kable_classic("hover", full_width = TRUE) %>%
    kableExtra::scroll_box(width = "100%")
    } else {
      ## get the minimum scan number
      minScan <- min(map_df(corPRT, length))
      ## trim scans
      corPRT2 <- map_df(corPRT, `[`, 1: minScan)
      ## correlation table
      TableMS2PRT <- as.data.frame(round(cosSim(corPRT2), 3)) %>%
        dplyr::mutate_all(~ kableExtra::cell_spec(.x, background = ifelse(.x < 0.85 , "#ff8080", "#FFFFFF"))) %>%
        kableExtra::kbl(escape = FALSE, 
                    caption = "Table 6: Precursor Distribution across RT similarity") %>%
        kableExtra::kable_classic("hover", full_width = TRUE) %>%
        kableExtra::scroll_box(width = "100%")
      }
}

```

```{r MS2Table, echo = FALSE, out.width = "100%"}

## use kableExtra instead of DT 
if(!is.null(getMS2)){
  kableExtra::kbl(getMS2, escape = FALSE, caption = "Table 4: Summary of MS2 events") %>%
    kableExtra::kable_classic("hover", full_width = TRUE)
} else {
  print("not exist")
}

```
 
## Precursor Distribution across Mass Range

`Precursor Distribution across Mass` plot is used to visualize the mass range proportion of fragmented precursors. 

- It can be used to check peaks at which mass ranges are mainly fragmented.

- High similarity in `Precursor Distribution across Mass` among QC samples indicates good reproducibility in MS2.

```{r MS2MZ, echo = FALSE, out.width = "100%"}

if(!is.null(MS2_mz)){
   MS2_mz 
} else {
  print("not exist")
}

```

`Cosine Similarity` is used to quantify the similarity of precursor distribution across mass range plots among QC samples. Cosine similarity over 0.85 indicate high similarity. If the value is below 0.85, it will be highlight in <span style = "color: #ff8080">**red**</span> in Table 5. 

> **Note that** precursor ions were binned by 10 Da for Pearson correlation calculation.

```{r MS2MZTable, echo = FALSE, out.width = "100%"}
if(!is.null(TableMS2Pmass)){
   TableMS2Pmass 
} else {
  print("not exist")
}

```


## Precursor Distribution across RT

`Precursor Distribution across RT` plot is used to visualize the number of fragmented precursors at each RT (or scan). It can be useful to spot any signal dropouts during data acquisition.
 
- In both **data-dependent acquisition (DDA)** and **data-independent acquisition (DIA)** mode, you are expected to see a continuous precursor ions distribution across the entire RT range.

- High similarity in `Precursor Distribution across RT` among QC samples indicates good reproducibility in MS2.


```{r MS2RT, echo = FALSE, out.width = "100%"}

if(!is.null(MS2_RT)){
   MS2_RT
} else {
  print("not exist")
}

```

`Cosine Similarity` is used to quantify the similarity of precursor distribution across RT plots among QC samples. Cosine similarity over 0.85 indicate high similarity. If the value is below 0.85, it will be highlight in <span style = "color: #ff8080">**red**</span> in Table 6. 

> **Note that** RT were binned by 0.05 min for Pearson correlation calculation.

```{r MS2RTTable, echo = FALSE, out.width = "100%"}
if(!is.null(TableMS2Pmass)){
   TableMS2PRT
} else {
  print("not exist")
}

```


# Reference

[1] Scalbert, A., Brennan, L., Fiehn, O., Hankemeier, T., Kristal, B.S., van Ommen, B., Pujos-Guillot, E., Verheij, E., Wishart, D. and Wopereis, S., 2009. Mass-spectrometry-based metabolomics: limitations and recommendations for future progress with particular focus on nutrition research. Metabolomics, 5(4), pp.435-458.

[2] Begou, O., Gika, H.G., Theodoridis, G.A. and Wilson, I.D., 2018. Quality Control and Validation Issues in LC-MS Metabolomics. Methods in molecular biology (Clifton, NJ), 1738, pp.15-26.
